Wired for Success: Sweden's Electrical Innovators in the Age of Industrial Transformation

### 2023-05-15 index.qmd
add emoji

---
title: "Wired for Success: Sweden's Electrical Innovators in the Age of Industrial Transformation"
title-slide-attributes:
  data-background-image: "assets/preview.png"
  data-background-size: "30%"
  data-background-opacity: "0.6"
format: 
  revealjs:
    code-overflow: wrap
    width: 1600
    height: 900
    fig-align: center
    default-image-extension: svg
    footer: "[{{< fa house >}}](https://j-jayes.github.io/Data-Visualization-in-R-Lund/)"
    slide-number: c
image: "preview_img/03-coordinate-systems-and-axes.jfif"
favicon: html/favicon.ico
---

@@

    - It made individuals richer
    - It lowered inequality
    - It allowed structural change on the labour market

@@

    - Who were the innovators or adopters?
    - Did they move to opportunity?
    - How much training did they have in the new technology?

@@

🧑
📝
🗑️

Arrow

💻🧹

Arrow

📚
🎓
💼

@@

- Many abbreviations

@@

## 1. Translate

@@

## 2. Structure

@@

## 3. Augment

@@

@@

@@

@@

## How are they easy now?

Now we can use the OCR / NLP OpenAI API to get a lot of information from the text of the patent.

>>>>

---
title: "Wired for Success: Sweden's Electrical Innovators in the Age of Industrial Transformation"
title-slide-attributes:
  data-background-image: "assets/preview.png"
  data-background-size: "30%"
  data-background-opacity: "0.6"
format: 
  revealjs:
    code-overflow: wrap
    width: 1600
    height: 900
    fig-align: center
    default-image-extension: svg
    footer: "[{{< fa house >}}](https://j-jayes.github.io/Data-Visualization-in-R-Lund/)"
    slide-number: c
    code-block-bg: true
    code-block-border-left: "#31BAE9"
image: "preview_img/03-coordinate-systems-and-axes.jfif"
favicon: html/favicon.ico
---

@@

    - It made individuals richer 💰🔋
    - It lowered inequality 📉🤝
    - It allowed structural change on the labor market 🔧💼

@@

    - Who were the innovators or adopters? 🤔🚀
    - Did they move to opportunity? 🏃‍♀️💼
    - How much training did they have in the new technology? 🎓⚡️

@@

## Data pipeline

| Step | Process                                   |        | Description                                                           |
|------|-------------------------------------------|-------------|
| 1    | Scrape book data from website             | 🌐➡️📚       | 
| 2    | Split biographies on each page of a book  | 📚➡️📄       | 
| **3**    | **Translate biographies**                     | 📄🔄📄       | 
| **4**    | **Structure biographies using Schema.org**    | 📄➡️🔗       | 
| **5**    | **Augment data**                              | 🔗🔍🌍🎓📚   |
| 6    | Store data                                | 🔗📚📂       |

@@

- Many abbreviations and contractions
    - DOB: "f\.\\s*(\\d{2})\\/(\\d{2})/(\\d{2})"
    - Gävleborg County: "Gävleb. l."

@@

## 1. Translate 🤖

### Using OpenAI's GPT-3.5 chat API:

We tell the system "You are an expert in Swedish family history"

Then we ask it to translate the Swedish text to English

@@

## 2. Structure 🤖

![](assets/schema.png)

## 2. Structure 🤖

### Using OpenAI's GPT-3.5 chat API:

We tell the system "You are an expert in Swedish family history and the Schema.org/Person format"

We provide it the original Swedish and English translation, and ask it to structure the biography in Schema.org/Person format as a JSON object

@@

## 3. Augment 🤖

### Using OpenAI's GPT-4 chat API:

We create python code to geocode any addresses in the biography (in this example), as well as flagging the highest level of education, and extract any mention of patents or inventions in the biography.

@@

TODO: fix names

@@

```{r}
library(ggtext)
library(showtext)
font_add_google("IBM Plex Mono", "ibm")
showtext_auto(enable = TRUE)

timeline <- tibble(
  event = c("Birth", "Graduated from Bergsskolan in Filipstad", 
            "Specialized studies at the Royal Institute of Technology (KTH)",
            "Studied at the Institute of Metallurgy and Stockholm University",
            "Marriage to Sigrid Johansson",
            "Birth of Ingvar", "Birth of Lennart",
            "Chemist at Strömsnäs Järnverks A-B in Degerfors",
            "Metallurgist and chemist at Westinghouse Electric & Manufacturing Co.",
            "Chief metallurgist at Laclede Steel Co. in Alton, Illinois, USA",
            "Furnace and steelworks engineer at A-B Iggesunds Bruk",
            "Site manager at Gunnebo Bruks Nya A-B, Varbergsverket",
            "Member of the municipal executive committee",
            "Deputy chairman of the economic department",
            "Deputy member of the board of the power plant",
            "Chairman of Varbergs Sparbank",
            "Employer representative in the district council of the county labor board",
            "Member of the board of Varbergs Luftskyddsförening",
            "Secretary of Varbergs Högerförening",
            "Chairman of the railway sick fund",
            "Plant Society for Small Bird Friends",
            "Traveled to Germany",
            "Traveled to Denmark and Czechoslovakia",
            "Traveled to Austria",
            "Traveled to the USA",
            "*Wrote 'Some fundamental factors for obtaining sharp thermal curves'",
            "*Wrote 'Contemporary production of saw blades, circular saws, and machine knives'",
            "Hobbies: hunting and fishing"
  ),
  start_year = c(1893, 1917, 1920, 1921, 1936, 1938, 1942, 1918, 1923, 1927, 1929,
                 1931, NA, NA, NA, NA, NA, NA, NA, NA, NA, 1921, 1921, 1921, 1923, 1925, 1931, NA),
  end_year = c(1893, 1917, 1922, 1922, NA, NA, NA, 1920, 1929, 1927, 1931,
               NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, 1923, 1923, 1921, 1929, NA, NA, NA)
)

timeline %>%
    filter(!is.na(start_year)) %>%
  arrange(start_year) %>%
  mutate(end_year_imputed = case_when(
    is.na(end_year) ~ lead(start_year),
    is.na(lead(start_year)) ~ start_year,
    TRUE ~ end_year
  )) %>%
  mutate(
    event = fct_reorder(event, start_year, .desc = TRUE)
  ) %>%
  # pivot_longer(c(start_year, end_year_imputed)) %>%
  # ggplot(aes(value, english_occupation, colour = name)) +
  ggplot(aes(xmin = start_year, xmax = end_year_imputed, y = event)) +
  geom_linerange(colour = "gray50", size = 2) +
  geom_point(aes(x = end_year_imputed), colour = "#be5d2b", size = 4) +
  geom_point(aes(x = start_year), colour = "#2B8CBE", size = 4) +
  labs(
    title = "Karl Lund's life course",
    subtitle = "Showing <span style = 'color:#2B8CBE;'>start dates</span> and <span style = 'color:#be5d2b;'>end dates</span> for his various roles and life events",
    x = NULL,
    y = NULL
  ) +
  theme(
    legend.position = "none",
    plot.subtitle = element_markdown(),
    plot.title.position = "plot",
    text = element_text(family = "ibm", size = 28)
  )

@@

Sample of birthplaces of individuals with surnames beginning with A in Götaland 1948:

@@

### How are they easy now?

Now we can use OCR / NLP to get a lot of information from the text of the patent.
